---
title: "Notebook: COVID Data"
author: "Nathan Hoteling"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

<hr>
<br>
<br>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(lubridate)
library(reshape2)
library(stringr)
library(RcppRoll)
library(tidyr)
library(readr)
library(plotly)
library(DT)
library(scales)

library(r2d3)
library(jsonlite)

fsave = TRUE

# colors:
# red (48% saturation):
red48 <- "#c0514d"
# red (60% saturation):
red60 <- "#cf443f"
# red (70% saturation):
red70 <- "#db3933"

orng1 <- "#e64d00"
```




```{r, include=FALSE}
####################
#  DATA URLs
####################
#
# JHU data:
# confirmed cases - WORLD
#
url_jhu_cfm <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"
#
# JHU data:
# deaths - WORLD
#
url_jhu_dth <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"
#
# JHU data:  NOT USED
# daily totals
# This looks like a subset of the data in the tables above...
#
url_jhu_dly <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-08-2020.csv"
#
# JHU data:
# confirmed cases - US STATES
#
url_jhu_uscfm <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv"
#
# JHU data:
# deaths - US STATES
#
url_jhu_usdth <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv"

#
#
# Vaccination data from Our World in Data (OWID)
# --> OWID pulls US vaccination data from CDC
# --> OWID pulls international vaccination data from different sources; 
#     url for each country is included in the data file
#
# US States
url_owd_usvac <- "https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/us_state_vaccinations.csv"
# World
url_owd_wdvac <- "https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations.csv"
#
#


#
# Other data available but not used in this analysis:
#
# Data from covidtracking.com -------------------------------------------------
url_cvt_dly <- "https://covidtracking.com/api/v1/states/daily.csv"
url_cvt_news <- "https://covidtracking.com/api/press"
#
#
# Data from NYTimes via GitHub ------------------------------------------------
# state data
#
url_nyt_ste <- "https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv"
# county data
url_nyt_cty <- "https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv"

```



```{r, include=FALSE}
# Note: fsave is set above
#
#
# READ DATA
#
#
df.cfmd <- read.table(url_jhu_cfm,sep=",",header=TRUE,fill=TRUE,quote="\"")  # jhu confirmed
df.deth <- read.table(url_jhu_dth,sep=",",header=TRUE,fill=TRUE,quote="\"")  # jhu deaths

df.uscfmd <- read.table(url_jhu_uscfm,sep=",",header=TRUE,fill=TRUE,quote="\"")  # jhu US confirmed
df.usdeth <- read.table(url_jhu_usdth,sep=",",header=TRUE,fill=TRUE,quote="\"")  # jhu US deaths
#df.daly <- read.table(url_jhu_dly,sep=",",header=TRUE,fill=TRUE,quote="\"")  # jhu daily xxx
#df.stat <- read.table(url_nyt_ste,sep=",",header=TRUE,fill=TRUE,quote="\"")  # nyt state
#df.cnty <- read.table(url_nyt_cty,sep=",",header=TRUE,fill=TRUE,quote="\"")  # nyt county


#
# Save the raw data
#
write_csv(df.cfmd, file="data/raw/jhu_cases.csv")
write_csv(df.deth, file="data/raw/jhu_deaths.csv")
#write_csv(df.daly, file="data/raw/jhu_daily.csv")
write_csv(df.uscfmd, file="data/raw/jhu_UScases.csv")
write_csv(df.usdeth, file="data/raw/jhu_USdeaths.csv")
#
# Vaccination data from OWID / CDC via GitHub
#
df.usvac <- read.table(url_owd_usvac, sep=",", header=TRUE,fill=TRUE,quote="\"") # cdc vacc
df.wdvac <- read.table(url_owd_wdvac, sep=",", header=TRUE,fill=TRUE,quote="\"") # world vacc
```



```{r, include=FALSE, warning=FALSE}
#
#
# DATA PROCESSING - WORLD
#
#
#

#
#
# World Population data
#
#
# World population data downloaded from World Bank:
# https://data.worldbank.org/indicator/SP.POP.TOTL
#
#fname_pop <- "data/API_SP.POP.TOTL_DS2_en_csv_v2_1976634/API_SP.POP.TOTL_DS2_en_csv_v2_1976634.csv"
df.pop <- read.table("data/country_populations.csv", sep=",",skip=4, header=TRUE, fill=TRUE) %>% 
  select(Country.Name, X2019) %>%
  rename(Country.Region = Country.Name, population = X2019)
df.tmp <- read.table("data/country_conversions.csv", header=TRUE, sep=",")
for (i in 1:nrow(df.tmp)) {
  df.pop$Country.Region[ df.pop$Country.Region == df.tmp$wb[i] ] <- df.tmp$jhu[i]
}



#
#
# World data from JHU
#
#
# confirmed cases ----
# there are columns for each date
# and in each column is the cumulative
# total number of cases
#

cases <- df.cfmd %>% 
  pivot_longer(cols = starts_with("X"), names_to = "date", values_to = "cases") %>%
  mutate(date = mdy(str_replace(date, pattern="X", replacement="")),
         key = paste(Province.State,Country.Region,date,sep="-"))

deaths <- df.deth %>%
  pivot_longer(cols = starts_with("X"), names_to = "date", values_to = "deaths") %>%
  mutate(date = mdy(str_replace(date, pattern="X", replacement="")),
         key = paste(Province.State,Country.Region,date,sep="-"))

jhu_combined <- cases %>% 
  full_join(deaths[,c(6,7)], by="key") %>% 
  select(-key) %>%
  group_by(Country.Region, date) %>% 
  summarise(cases=sum(cases), deaths=sum(deaths)) %>%
  left_join(df.pop, by="Country.Region")

countries <- unique(jhu_combined$Country.Region)
ncountries <- length(countries)
d <- lapply(seq_len(ncountries), function(i) {
  print(paste(i, countries[i]))
  df <- jhu_combined %>% filter(Country.Region == countries[i]) %>% arrange(date)
  df$cases1 <- df$cases - dplyr::lag(df$cases)        # daily cases
  df$deaths1 <- df$deaths - dplyr::lag(df$deaths)     # daily deaths
  df$cases7 <- roll_meanr(df$cases1, n=7)              # 7-day mean cases
  df$deaths7 <- roll_meanr(df$deaths1, n=7)            # 7-day mean deaths
  df$Cases1 <- df$cases1/df$population
  df$Deaths1 <- df$deaths1/df$population
  df$Cases7 <- df$cases7/df$population
  df$Deaths7 <- df$deaths7/df$population
  return(df)
})
jhu <- do.call(rbind, d)

#
#  Filter out some countries so the resulting
#  plots aren't quite so messy
#
trim <- 50  # keep only countries with top n population
pop_trim <- jhu %>% 
  group_by(Country.Region) %>% 
  summarise(pop = max(population)) %>% 
  arrange(desc(pop)) %>% 
  top_n(trim)

jhu_trim <- jhu %>% filter(Country.Region %in% pop_trim$Country.Region)


#
#
# Save the processed data
# jhu: World data
# jhu_us: State data
#
if (fsave) { 
  write_rds(jhu, file=paste("data/rds/covid-jhu-",Sys.Date(),".rds",sep="")) 
  #write_rds(jhu_us, file=paste("data/rds/covid-jhu-us-",Sys.Date(),".rds",sep=""))
  }
```



```{r, include=FALSE, warning=FALSE}
#
#
# DATA PROCESSING - STATES
#
#
#

#
# US population data downloaded from Census
#
df.census <- read.table("data/nst-est2020.csv", sep=",", header=TRUE, fill=TRUE) %>%
  rename(region = REGION,
         division = DIVISION,
         id = STATE,
         state = NAME,
         population = POPESTIMATE2020) %>%
  select(region, division, id, state, population)

#
#
# US States data from JHU
#
#
cases <- df.uscfmd %>% 
  pivot_longer(cols = starts_with("X"), names_to = "date", values_to = "cases") %>%
  mutate(date = mdy(str_replace(date, pattern="X", replacement=""))) %>%
  group_by(Province_State, date) %>%
  summarise(cases = sum(cases))

deaths <- df.usdeth %>%
  pivot_longer(cols = starts_with("X"), names_to = "date", values_to = "deaths") %>%
  mutate(date = mdy(str_replace(date, pattern="X", replacement=""))) %>%
  group_by(Province_State, date) %>%
  summarise(deaths = sum(deaths))

jhu_uscombined <- cases %>% 
  full_join(deaths, by=c("Province_State","date")) %>% 
  rename(state = Province_State) %>%
  left_join(df.census, by="state")

states <- unique(jhu_uscombined$state)
nstates <- length(states)
d <- lapply(seq_len(nstates), function(i) {
  print(paste(i, states[i]))
  df <- jhu_uscombined %>% filter(state == states[i]) %>% arrange(date)
  df$cases1 <- df$cases - dplyr::lag(df$cases)        # daily cases
  df$deaths1 <- df$deaths - dplyr::lag(df$deaths)     # daily deaths
  df$cases7 <- roll_meanr(df$cases1, n=7)              # 7-day mean cases
  df$deaths7 <- roll_meanr(df$deaths1, n=7)            # 7-day mean deaths
  df$Cases1 <- df$cases1/df$population
  df$Deaths1 <- df$deaths1/df$population
  df$Cases7 <- df$cases7/df$population
  df$Deaths7 <- df$deaths7/df$population
  return(df)
})
jhu_us <- do.call(rbind, d)

census_regions <- data.frame(region=c(1,2,3,4),
                             name=c("Northeast","Midwest","South","West"))
census_divisions <- data.frame(division=c(1,2,3,4,5,6,7,8,9),
                               name=c("New England", "Middle Atlantic","East North Central",
                                      "West North Central", "South Atlantic","East South Central",
                                      "West South Central","Mountain","Pacific"))
jhu_us$regionName <- census_regions$name[ match(jhu_us$region, census_regions[,1]) ]
jhu_us$divisionName <- census_divisions$name[ match(jhu_us$division, census_divisions[,1]) ]


#
# Vaccination Data
#
#
df.vaxUS <- df.usvac %>%
  select(date, location, people_fully_vaccinated, people_fully_vaccinated_per_hundred) %>%
  mutate(date = as.Date(date)) %>%
  rename(state = location, 
         Vaxd = people_fully_vaccinated_per_hundred,
         vaxd = people_fully_vaccinated) %>%
  left_join(df.census, by="state") %>%
  mutate(Vax = 100*vaxd/population)
df.vaxUS$regionName <- census_regions$name[ match(df.vaxUS$region, census_regions[,1]) ]
df.vaxUS$divisionName <- census_divisions$name[ match(df.vaxUS$division, census_divisions[,1]) ]
df.vaxUS <- df.vaxUS %>% filter(!is.na(regionName))
#
#
# Save the processed data
# jhu: World data
# jhu_us: State data
#
if (fsave) { 
  #write_rds(jhu, file=paste("data/rds/covid-jhu-",Sys.Date(),".rds",sep="")) 
  write_rds(jhu_us, file=paste("data/rds/covid-jhu-us-",Sys.Date(),".rds",sep=""))
  }
```





```{r, include=FALSE, warning=FALSE, echo=FALSE}
#
#  PLOTS - WORLD Data
#
#


#selected_countries1 <- c("US", "Spain", "Italy", "United Kingdom", 
#                        "Germany", "France", "Iran", "Canada", 
#                        "India", "Ireland", "Austria", "Japan", 
#                        "Korea, South", "China")
selected_countries1 <- c("Spain", "Italy", "United Kingdom", 
                        "Germany", "Brazil", "Iran","Sweden",  
                        "Ireland", "Japan", "China")

jhu_usa <- jhu_trim %>% filter(Country.Region %in% c("US"))

x_buffer <- difftime(max(jhu_trim$date), min(jhu_trim$date), unit="weeks")*0.20
df.world <- jhu_trim %>%
  mutate(Cases7 = 1e5*Cases7,
         Deaths7 = 1e5*Deaths7) %>%
  rename(Country = Country.Region)

df.usa <- jhu_usa %>%
  mutate(Cases7 = 1e5*Cases7,
         Deaths7 = 1e5*Deaths7) %>%
  rename(Country = Country.Region)

#
# Daily cases per 100k, 7-day rolling average
#
p.Cases <- ggplot(df.world, aes(x=date,y=Cases7,group=Country)) +
  geom_line(aes(text = paste(" Date:", format.Date(date, "%m/%d/%Y"), "<br>",
                             "Country:", Country, "<br>",
                             "Cases:", round(Cases7, digits=1), "per 100k"), sep=""),
            color="grey50",size=0.2, alpha=0.5) +
  geom_line(data=df.usa, color="grey99",size=1.2, alpha=1.0) +
  geom_line(data=df.usa, aes(text = paste(" Date:", format.Date(date, "%m/%d/%Y"), "<br>",
                             "Country:", Country, "<br>",
                             "Cases:", round(Cases7, digits=1), "per 100k"), sep=""), 
            color=orng1,size=0.8, alpha=0.8) +
  geom_point(data=df.usa %>% filter(date == max(date)),
             aes(x=date,y=Cases7),
             pch=21, size=3, color="grey99",fill=orng1) +
  #geom_text(data=df.usa %>% filter(date == max(date)), aes(label=Country, text=""),
  #          color="grey99", size=5,hjust=-0.5, fontface="bold") +
  geom_text(data=df.usa %>% filter(date == max(date)), 
            aes(x=date+days(4), label=Country),
            color=orng1, size=5,hjust=-0.2) +
  labs(x="",y="", title = "Worldwide COVID cases per 100k population") +
  scale_x_date(date_labels = "%b", limits=c(min(jhu_trim$date), max(jhu_trim$date)+days(7))) +
  lims(y = c(0,1.2*max(df.world$Cases7))) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text = element_text(size=10))


#
# Daily deaths per 100k, 7-day rolling average
#
p.Deaths <- ggplot(df.world, aes(x=date,y=Deaths7,group=Country)) +
  geom_line(aes(text = paste(" Date:", format.Date(date, "%m/%d/%Y"), "<br>",
                             "Country:", Country, "<br>",
                             "Deaths:", round(Deaths7, digits=1), "per 100k"), sep=""),
            color="grey50",size=0.2, alpha=0.5) +
  geom_line(data=df.usa, color="grey99",size=1.2, alpha=1.0) +
  geom_line(data=df.usa, aes(text = paste(" Date:", format.Date(date, "%m/%d/%Y"), "<br>",
                             "Country:", Country, "<br>",
                             "Deaths:", round(Deaths7, digits=1), "per 100k"), sep=""),
            color=orng1,size=0.8, alpha=0.8) +
  geom_point(data=df.usa %>% filter(date == max(date)),
             pch=21, size=3, color="grey99",fill=orng1) +
  #geom_text(data=jhu_usa %>% filter(date == max(date)),
  #          aes(label=Country.Region),
  #          color="grey99", size=5,hjust=-0.2, fontface="bold") +
  geom_text(data=df.usa %>% filter(date == max(date)),
            aes(x=date+days(4), label=Country),
            color=orng1, size=5,hjust=-0.2) +
  labs(x="Date",y="", title="Worldwide COVID deaths per 100k population") +
  scale_x_date(date_labels = "%b", limits=c(min(jhu_trim$date), max(jhu_trim$date)+days(7))) +
  lims(y = c(0,1.2*max(jhu_trim$Deaths7))) +
  #theme_nh
  theme_minimal() +
  theme(legend.position = "none")
```


## World data {.tabset}

The per capita numbers of cases and deaths for countries with the top 50 world populations are listed here.  The table shows the 7-day averages for the most recent date.

### Cases

```{r, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=5}
# Draw plotly versions of the plots
ggplotly(p.Cases, tooltip=c("text")) %>% style(textposition = "right")
```

### Deaths

```{r, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=5}
# Draw plotly versions of the plots
ggplotly(p.Deaths, tooltip=c("text")) %>% style(textposition = "right")
```

### Table

```{r, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
df.w <- df.world %>% 
  filter(date == max(date)) %>%
  select(Country, Cases7, Deaths7,cases7,deaths7) %>%
  arrange(desc(Cases7))
DT::datatable(df.w,
              colnames = c("Country", "Cases per 100k", 
                           "Deaths per 100k", "Cases", "Deaths")) %>%
  formatRound(2,1) %>%
  formatRound(3,2) %>%
  formatRound(4:5,0)
```





```{r, include=FALSE, warning=FALSE, echo=FALSE, fig.width=8}
#
#  PLOTS - STATE Data
#
#

x_buffer <- difftime(max(jhu_us$date), min(jhu_trim$date), unit="weeks")*0.20
df.us <- jhu_us %>% filter(!is.na(regionName)) %>%
  mutate(Cases7 = 1e5*Cases7,
         Deaths7 = 1e5*Deaths7)

df.va <- df.us %>% filter(state=="Virginia")


#
# Daily cases per 100k, 7-day rolling average
#
p.CasesUS <- ggplot(df.us, aes(x=date,y=Cases7,group=state)) +
  geom_line(aes(text = paste(" Date:", format.Date(date, "%m/%d/%Y"), "<br>",
                             "State:", state, "<br>",
                             "Cases:", round(Cases7, digits=1), "per 100k"), sep=""),
            color="grey50",size=0.2, alpha=0.5) +
  geom_line(data=df.va, color="grey99",size=1.2, alpha=1.0) +
  geom_line(data=df.va, aes(text = paste(" Date:", format.Date(date, "%m/%d/%Y"), "<br>",
                             "State:", state, "<br>",
                             "Cases:", round(Cases7, digits=1), "per 100k"), sep=""), 
            color=orng1,size=0.8, alpha=0.8) +
  geom_point(data=df.va %>% filter(date == max(date)),
             aes(x=date,y=Cases7),
             pch=21, size=3, color="grey99",fill=orng1) +
  #geom_text(data=df.usa %>% filter(date == max(date)), aes(label=Country, text=""),
  #          color="grey99", size=5,hjust=-0.5, fontface="bold") +
  geom_text(data=df.va %>% filter(date == max(date)), 
            aes(x=date+days(4), label="VA"),
            color=orng1, size=5,hjust=-0.2) +
  labs(x="",y="", title = "US COVID cases per 100k population") +
  scale_x_date(date_labels = "%b", breaks=date_breaks("2 months"),
               limits=c(min(jhu_trim$date), max(jhu_trim$date)+days(7))) +
  lims(y = c(0,1.2*max(jhu_trim$Cases7))) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text = element_text(size=10))


#
# Daily deaths per 100k, 7-day rolling average
#
p.DeathsUS <- ggplot(df.us, aes(x=date,y=Deaths7,group=state)) +
  geom_line(aes(text = paste(" Date:", format.Date(date, "%m/%d/%Y"), "<br>",
                             "State:", state, "<br>",
                             "Deaths:", round(Deaths7, digits=1), "per 100k"), sep=""),
            color="grey50",size=0.2, alpha=0.5) +
  geom_line(data=df.va, color="grey99",size=1.2, alpha=1.0) +
  geom_line(data=df.va, aes(text = paste(" Date:", format.Date(date, "%m/%d/%Y"), "<br>",
                             "State:", state, "<br>",
                             "Deaths:", round(Deaths7, digits=1), "per 100k"), sep=""), 
            color=orng1,size=0.8, alpha=0.8) +
  geom_point(data=df.va %>% filter(date == max(date)),
             aes(x=date,y=Deaths7),
             pch=21, size=3, color="grey99",fill=orng1) +
  #geom_text(data=df.usa %>% filter(date == max(date)), aes(label=Country, text=""),
  #          color="grey99", size=5,hjust=-0.5, fontface="bold") +
  geom_text(data=df.va %>% filter(date == max(date)), 
            aes(x=date+days(4), label="VA"),
            color=orng1, size=5,hjust=-0.2) +
  labs(x="",y="", title = "US COVID deaths per 100k population") +
  scale_x_date(date_labels = "%b", limits=c(min(jhu_trim$date), max(jhu_trim$date)+days(7))) +
  lims(y = c(0,1.2*max(jhu_trim$Deaths7))) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text = element_text(size=10))
```

<br>
<br>

## USA data {.tabset}

The per capita numbers of cases and deaths for all states are listed here.  The table shows the 7-day averages for the most recent date.

### Cases

```{r, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4}
ggplotly(p.CasesUS, tooltip=c("text")) %>% style(textposition = "right")
```

### Deaths

```{r, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4}
ggplotly(p.DeathsUS, tooltip=c("text")) %>% style(textposition = "right")
```

### Table

```{r, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
df.u <- df.us %>% 
  filter(date == max(date)) %>%
  select(state, regionName, Cases7, Deaths7,cases7,deaths7) %>%
  arrange(desc(Cases7))
DT::datatable(df.u,
              colnames = c("State", "Region", "Cases per 100k", 
                           "Deaths per 100k", "Cases", "Deaths")) %>%
  formatRound(3,1) %>%
  formatRound(4,2) %>%
  formatRound(5:6,0)
```



```{r, include=FALSE, warning=FALSE, echo=FALSE, fig.width=8}
#
#  PLOTS - STATE Data
#
#


#
# Daily cases per 100k, 7-day rolling average
#
cols1 <- c("South"="#fbb4ae","West"="#b3cde3","Northeast"="#ccebc5","Midwest"="#decbe4")
cols2 <- c("South"="#66c2a5","West"="#fc8d62","Northeast"="#8da0cb","Midwest"="#e78ac3")




# Colored by Census Region
p.CasesUS_misc <- ggplot(df.us, aes(x=date,y=Cases7,group=state)) +
  geom_line(aes(color=regionName),size=0.3, alpha=0.7) +
  scale_x_date(date_labels = "%b", breaks=date_breaks("2 months"),
               limits=c(min(jhu_trim$date), max(jhu_trim$date))) +
  scale_color_manual(name="Region",values=cols2) +
  labs(x="",y="", title = "") +
  lims(y = c(0,1.2*1e5*max(jhu_trim$Cases7))) +
  #facet_wrap(~regionName, ncol=2) +
  theme_minimal() +
  theme(legend.position = c(0.1,0.85),
        legend.background = element_rect(fill="grey100",
                                  size=0.2, linetype="solid", 
                                  colour ="grey75"),
        strip.text = element_text(size=12, face="bold", color="grey20"),
        axis.text = element_text(size=10))


# Colored by Census Region
p.DeathsUS_misc <- ggplot(df.us, aes(x=date,y=Deaths7,group=state)) +
  geom_line(aes(color=regionName),size=0.3, alpha=0.7) +
  scale_x_date(date_labels = "%b", breaks=date_breaks("2 months"),
               limits=c(min(df.us$date), max(df.us$date))) +
  scale_color_manual(name="Region",values=cols2) +
  labs(x="",y="", title = "") +
  lims(y = c(0,1.2*1e5*max(df.us$Deaths7))) +
  #facet_wrap(~regionName, ncol=2) +
  theme_minimal() +
  theme(legend.position = c(0.1,0.85),
        legend.background = element_rect(fill="grey100",
                                  size=0.2, linetype="solid", 
                                  colour ="grey75"),
        strip.text = element_text(size=12, face="bold", color="grey20"),
        axis.text = element_text(size=10))

p.vax <- ggplot(df.vaxUS, aes(x=date, y=Vax, group=state)) +
  geom_line(aes(color=regionName),size=0.3, alpha=0.7) +
  scale_x_date(date_labels = "%b", breaks=date_breaks("2 months")  ) +#,
               #limits=c(min(df.us$date), max(df.us$date))) +
  scale_color_manual(name="Region",values=cols2) +
  labs(x="", y="Percent Fully Vaccinated") +
  lims(y=c(0,100)) +
  theme_minimal() +
  theme(legend.position = c(0.1,0.80),
        legend.background = element_rect(fill="grey100",
                                  size=0.2, linetype="solid", 
                                  colour ="grey75"),
        strip.text = element_text(size=12, face="bold", color="grey20"),
        axis.text = element_text(size=10))
```

<br>
<br>

## Miscellaneous Plots

<details>
<summary style="size: 18px"> <b>US Cases by Census Region</b> </summary>
```{r, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4}
p.CasesUS_misc
```
</details>

<details>
<summary style="size: 18px"> <b>US Deaths by Census Region</b> </summary>
```{r, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4}
p.DeathsUS_misc
```
</details>

<details>
<summary style="size: 18px"> <b>US Vaccinations by Census Region</b> </summary>
```{r, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4}
p.vax
```
</details>


```{r, include=FALSE, warning=FALSE}
#####################
#                   #
# Data for D3 plots #
#                   #
#####################

# height & width for SVG
h <- 400
w <- 950


## World data
df.jhu <- jhu_trim %>% 
  rename(country = Country.Region) %>%
  mutate(cases = ifelse(is.na(cases), 0, cases),
         cases1 = ifelse(is.na(cases1), 0, cases1),
         cases7 = ifelse(is.na(cases7), 0, cases7),
         Cases1 = ifelse(is.na(Cases1), 0, 1e5*Cases1),
         Cases7 = ifelse(is.na(Cases7), 0, 1e5*Cases7),
         deaths = ifelse(is.na(deaths), 0, deaths),
         deaths1 = ifelse(is.na(deaths1), 0, deaths1),
         deaths7 = ifelse(is.na(deaths7), 0, deaths7),
         Deaths1 = ifelse(is.na(Deaths1), 0, 1e5*Deaths1),
         Deaths7 = ifelse(is.na(Deaths7), 0, 1e5*Deaths7))

## State data
df.usjhu <- jhu_us %>% 
  mutate(cases = ifelse(is.na(cases), 0, cases),
         cases1 = ifelse(is.na(cases1), 0, cases1),
         cases7 = ifelse(is.na(cases7), 0, cases7),
         Cases1 = ifelse(is.na(Cases1), 0, 1e5*Cases1),
         Cases7 = ifelse(is.na(Cases7), 0, 1e5*Cases7),
         deaths = ifelse(is.na(deaths), 0, deaths),
         deaths1 = ifelse(is.na(deaths1), 0, deaths1),
         deaths7 = ifelse(is.na(deaths7), 0, deaths7),
         Deaths1 = ifelse(is.na(Deaths1), 0, 1e5*Deaths1),
         Deaths7 = ifelse(is.na(Deaths7), 0, 1e5*Deaths7))


d.jhu <- jsonlite::toJSON(df.jhu)
d.usjhu <- jsonlite::toJSON(df.usjhu)
# Write json file for my pure d3.js version
write(d.jhu,"data/rcovid_world.json")
write(d.usjhu, "data/rcovid_state.json")
```







```{r, eval=FALSE, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
#
#

r2d3::r2d3(data=jsonlite::read_json("data/rcovid_world.json"),
     dependencies=list("js/d3-tip.js", "js/d3-legend.js"), 
     css="css/styles4.css",
     script="js/covid-world-cases.js",
     height = h,
     width = w,
     d3_version = 4,
     elementId = "covid1")
```





```{r, eval=FALSE, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
#
#

r2d3::r2d3(data=jsonlite::read_json("data/rcovid_world.json"),
     dependencies=list("js/d3-tip.js", "js/d3-legend.js"), 
     css="css/styles4.css",
     script="js/covid-world-deaths.js",
     height = h,
     width = w,
     d3_version = 4)

# height = 900
# width = 800
```

<br>
<br>

## Data Sources

COVID data on cases & deaths obtained from the following sources:  
- Johns Hopkins [GitHub site](https://github.com/CSSEGISandData/COVID-19)  
- Data from [covidtracking website](https://covidtracking.com/)  
- NY Times [GitHub site](https://github.com/nytimes/covid-19-data)  

COVID vaccine data from [Our World in Data](https://ourworldindata.org/covid-vaccinations):  
- Our World in Data posts regular updates on their [GitHub site](https://github.com/owid/covid-19-data/tree/master/public/data)  
- OWID data for US vaccines are pulled from [CDC Vaccine Data Tracker](https://covid.cdc.gov/covid-data-tracker/#vaccinations)  
- OWID data for international locations aggregated from a variety of sources, as documented on their [GitHub page](https://github.com/owid/covid-19-data/tree/master/public/data)  

Worldwide Population data from [World Bank](https://data.worldbank.org/indicator/SP.POP.TOTL)  
US Population data from [Census Website](https://www2.census.gov/programs-surveys/popest/datasets/2010-2020/national/totals/)  

<br>
<br>